WSClean: a new implementation of a fast, generic wide-field imager

### 2014-01-22 articles/2014/Article-WSClean/wsclean.tex
Progress

TODO: Algorithm; plain w-stacking and further alterations such as w-snapshots and full-sky imaging.

@@

\begin{table}[h]
 \caption{Scaling of the computational cost for various operations} \label{tbl:computational-cost-per-operation}
 \begin{tabular}{lll}
   \textbf{Operation} & \textbf{W-stacking} & \textbf{W-projection} \\
   Fourier transform(s) & $N_\textrm{wlay} N^2_\textrm{pix} \log N_\textrm{pix}$ & $N^2_\textrm{pix} \log N_\textrm{pix}$ \\
   W-term corrections   & $N_\textrm{wlay} N^2_\textrm{pix}$ & $N_\textrm{vis} N^2_\textrm{wkern}$ \\
   Gridding & $N_\textrm{vis}N^2_\textrm{kern}$ & $N_\textrm{vis} N_\textrm{kern}^2$
 \end{tabular}
\end{table}

Table~\ref{tbl:computational-cost-per-operation} shows how the computational costs scale for the operations that dominate the imaging in the w-stacking and w-projection algorithms. For comparison, we can assume the gridding kernel adds a constant term and the terms $N_\textrm{wlay}$ and $N_\textrm{wkern}$ follow approximately $w_e \sin \alpha_\textrm{FOV}$. The time complexity for w-stacking is then given by
\begin{equation}
\textrm{TC}_\textrm{wstacking}=\mathcal{O}\left(N^2_\textrm{pix} \log N_\textrm{pix} \sin \alpha_\textrm{FOV}+ N_\textrm{vis} \right),
\end{equation}
and for $w$-projection it is
\begin{equation}
\textrm{TC}_\textrm{wprojection}=\mathcal{O}\left(N^2_\textrm{pix} \log N_\textrm{pix} + N_\textrm{vis} w_e^2 \sin^2 \alpha_\textrm{FOV}\right).
\end{equation}
From these bounds it can be concluded that in the limiting behaviour, the $w$-stacking method will be faster when the gridding of the visibilities is the dominating cost of the algorithm. The $w$-projection algorithm will be faster when the inverse FFTs are dominating. In Sect.~\ref{sec:analysis} we will determine which method is faster in practice for different parameters.

@@

\section{The WSClean imager implementation} \label{sec:implementation}
We have implemented the $w$-stacking algorithm in the C++ language. Several steps are multi-threaded in this implementation: reading and writing, gridding from and to different $w$-layers, performing the Fast Fourier transforms and peak-finding and image subtraction during Cleaning. In addition, intrinsics were used for the latter. This speeds up a minor Cleaning iterations,

>>>>

@@

\begin{table}[h]
 \caption{Scaling of the computational cost for various imaging steps} \label{tbl:computational-cost-per-operation}
 \begin{tabular}{lll}
   \textbf{Operation} & \textbf{W-stacking} & \textbf{W-projection} \\
   \hline\hline
   Fourier transform(s) & $N_\textrm{wlay} N^2_\textrm{pix} \log N_\textrm{pix}$ & $N^2_\textrm{pix} \log N_\textrm{pix}$ \\
   W-term corrections   & $N_\textrm{wlay} N^2_\textrm{pix}$ & $N_\textrm{vis} N^2_\textrm{wkern}$ \\
   Gridding & $N_\textrm{vis}N^2_\textrm{kern}$ & $N_\textrm{vis} N_\textrm{kern}^2$ \\
   \hline\hline
 \end{tabular}
\end{table}

Table~\ref{tbl:computational-cost-per-operation} shows how the computational costs scale for the operations that dominate the imaging in the w-stacking and w-projection algorithms. For comparison, we can assume the gridding kernel adds a constant term and the terms $N_\textrm{wlay}$ and $N_\textrm{wkern}$ follow approximately $w_e \sin \alpha_\textrm{FOV}$. The time complexity for w-stacking is then given by
\begin{equation}
\textrm{TC}_\textrm{wstacking}=\mathcal{O}\left(N^2_\textrm{pix} \log N_\textrm{pix} w_e \sin \alpha_\textrm{FOV}+ N_\textrm{vis} \right),
\end{equation}
and for $w$-projection it is
\begin{equation}
\textrm{TC}_\textrm{wprojection}=\mathcal{O}\left(N^2_\textrm{pix} \log N_\textrm{pix} + N_\textrm{vis} w_e^2 \sin^2 \alpha_\textrm{FOV}\right).
\end{equation}
From these bounds it can be concluded that in the limiting behaviour, the $w$-stacking method will be faster when the gridding of the visibilities is the dominating cost of the algorithm. The $w$-projection algorithm will be faster when the inverse FFTs are dominating. In Sect.~\ref{sec:analysis} we will determine which method is faster in practice for different parameters.

\subsection{Full-sky imaging}
For low-frequency telescopes, it can be very useful to do full-sky imaging. In the case of MWA, the tile beam can have strong sidelobes at a distance of more than 90 degrees from the pointing centre. Imaging these sidelobes can be useful because there might be science (e.g. when searching for transients), or for self-calibration or full-sky deconvolution. The latter is for example useful when a strong source in a sidelobe affects the sensitivity in the pointing centre.

@@

\section{The WSClean imager implementation} \label{sec:implementation}
We have implemented the $w$-stacking algorithm in the C++ language. Several steps are multi-threaded in this implementation: reading and writing, gridding from and to different $w$-layers, performing the Fast Fourier transforms and peak-finding and image subtraction during Cleaning. For the latter, intrinsics were used as well. This speeds up a minor Cleaning iterations,
